Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-05-18 Chapter-5/5-Preparatory.tex
s

\section{Energy Calibration and Studies}\label{ch:energyCal}

>>>>

\clearpage

\section{Energy Calibration and Studies}\label{ch:energyCal} 
\subsection{Energy Calibration}\label{ch:energyCalibration}
\subsection{Uncertainty on Kinetic Energy}\label{ch:kinEn}
The measured kinetic energy of a hadron candidate at each argon slab determines which bins of the interacting and incident histograms a selected event is going to fill. With this study, we determine the uncertainty of the kinetic energy measurement which we will propagate into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:SysUncertaintyXSRawKaon} for the kaon cross section.

The kinetic energy of a hadron at the $j^{\text{th}}$ slice of argon in the TPC is given by

\begin{equation}
KE_{j} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam}^2 - E_{Loss} - E_{\text{FF-j}},
\end{equation}

where $p_{Beam}$ is the momentum measured by the beamline detectors,  $m_{Beam}$ is the mass of the hadron as reported in the PDG,  $E_{Loss}$  is the energy loss between the beamline and the TPC, and $ E_{\text{FF-j}}$ is the energy that the hadron deposited from the TPC front face until the $j^{\text{th}}$ slice.
The uncertainty on $KE_{j}$ is then given by 
\begin{equation}
\delta KE_{j} = \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2},
\end{equation}

where we have dropped the uncertainty on the mass, since it is orders of magnitude smaller than the other uncertainties.
We assume the relative uncertainty on $p_{Beam}$ to be 2\%, and the uncertainty on the energy loss upstream to be 7~MeV, as calculated in Section \ref{ch:eloss}. We describe the estimate of the uncertainty on $E_{\text{FF-j}}$ in the rest of this section.

The energy deposited from the TPC front face until the $j^{\text{th}}$ slice is the sum of the measured energy deposited in each previous slabs $E_{i}$, i.e.
\begin{equation}
E_{\text{FF-j}} = \sum_{i<j} E_{i}, 
\end{equation}
where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch, $Pitch_i$, for that point. 
If we assume conservatively that the measurements of $E_{i}$ are not independent from one another, the uncertainty on $E_{\text{FF-j}}$ becomes
\begin{equation}
\delta E_{\text{FF-j}} = (j-1) \delta E_{i}, 
\end{equation}
where $\delta E_{i}$ is the uncertainty on the energy loss in one slab of argon.

Figure \ref{fig:EnergyDeposited} shows the distribution of the energy deposited in each slab of argon, for the 60A negative pion dataset in black and for the pion only MC in blue. The distributions are fitted with a landau displayed in red for data and in teal for MC.
The uncertainty on $E_{i}$ is given by the width of the Landau fit to the data. A small systematic uncertainty  is given by a 1.0\% difference between the most probable value of the landau fits in data and MC.

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{Chapter-5/Images/DepEnergy_Fit_v4.png}
\caption[]{ Energy Deposited in Pion MC and 60A data.  } \label{fig:EnergyDeposited}
\end{figure}

So the uncertainty on the incident kinetic energy is given by

\begin{equation}
\delta KE^{Incident} = \sqrt{ (\delta KE_{Initial})^2 + (\delta E^{Slab}_{j})^2} = \sqrt{(12 MeV)^2 + (2 MeV)^2} = 12.1 MeV
\end{equation}

Figure \ref{fig:EnergyDepositedStacked} shows the stacked version of the Energy Deposited plots with the backgrounds stacked. The backgrounds are given in the ratio of 68.8\% pion, 4.6\% muon, and 26.6\% electron. Once they are taken in these ratios, the sum of the MC is normalized to the sum of the data.

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DepEnergy_Stacked_v1.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DepEnergy_Stacked_v3.png}
\caption[]{ Energy Deposited with all the MC and 60A data.  } \label{fig:EnergyDepositedStacked}
\end{figure}

The energy at the interacting point is given by
\begin{equation}
KE_{Interaction} = \sqrt{P_{WCtrk}^2 + m_{\pi}^2} - E_{Loss} - (\Sigma dE/dX_{i} \times Pitch)
\end{equation}

and has the exact same uncertainty as the incident kinetic energy plot. Thus these estimates can be applied to getting the uncertainty on the energy of the reconstructed cross-section.

A study we did was to look at the difference between DATA/MC in the dE/dX and energy deposited. We basically found there is very little difference between the two and we try to quantify how much the difference is.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{dE/dX}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Figure \ref{fig:dEdXLinearScale} shows the output of the fit of the Pion MC and the 60 Amp data. The MC is normalized to the data and both are fit to a Landau function. \footnote{The entries at dE/dX = 0 come from an uninitialized variable and can/should be taken out of these plots}

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v1.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v4.png}
\caption[]{ dE/dX for 60Amp data and data driven pion MC, both fit with a Landau  } \label{fig:dEdXLinearScale}
\end{figure}

The difference between the two MPV's, is 2.4\% between the data and the MC.

Figure \ref{fig:dEdXLinearStacked} shows the stacked version of the dE/dX with the backgrounds stacked. The backgrounds are given in the ratio of 68.8\% pion, 4.6\% muon, and 26.6\% electron. Once they are taken in these ratios, the sum of the MC is normalized to the sum of the data.

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v1.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v4.png}
\caption[]{ Stacked versions of the dE/dX with the data and electron/muon/pion MC.  } \label{fig:dEdXLinearStacked}
\end{figure}

For completeness, the log scale versions of are shown in Figure \ref{fig:dEdXLogScale}.

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v2.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v3.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v2.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v3.png}
\caption[]{ dE/dX for 60Amp data and MC shown in log scale  } \label{fig:dEdXLogScale}
\end{figure}

Plotting scripts can be found here on lariatgpvm \begin{verbatim}
/lariat/app/users/jasaadi/v06_34_01_PionWeek/PlottingScripts
\end{verbatim} and the samples were put here \begin{verbatim}
/lariat/data/users/elenag/theFinalPions/TPCDATA

/lariat/data/users/elenag/theFinalPions/TPC_MC/
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Energy Deposited}\label{sec:Energy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
